\chapter{Compiling XCSoar}\label{cha:compiling}

The \texttt{make} command is used to launch the XCSoar build process.
You can learn more about the build system internals in chapter
\ref{cha:build}.

Most of this chapter describes how to build XCSoar on Linux, with
examples for Debian/Ubuntu.  A cross-compiler is used to build
binaries for other operating systems (for example Android and
Windows).

\section{Getting the Source Code}

The XCSoar source code is managed with
\href{http://git-scm.com/}{git}.  It can be downloaded with the
following command:

\begin{verbatim*}
git clone git://github.com/XCSoar/XCSoar
\end{verbatim*}

To update your repository, type:

\begin{verbatim*}
git pull
\end{verbatim*}

To update third-party libraries used by XCSoar (such as
\href{http://www.boost.org/}{Boost}), type:

\begin{verbatim*}
git submodule init
git submodule update
\end{verbatim*}

For more information, please read to the git documentation.

\section{Requirements}

The following is needed for all targets:

\begin{itemize}
\item GNU make
\item GNU compiler collection (\texttt{gcc}), version 5 or later
  or clang/LLVM 3.8 (with "make CLANG=y")
\item GNU gettext
\item \href{http://librsvg.sourceforge.net/)}{rsvg}
\item \href{http://www.imagemagick.org/}{ImageMagick 6.4}
\item \href{http://xmlsoft.org/XSLT/xsltproc2.html}{xsltproc}
\item \href{http://www.info-zip.org/}{Info-ZIP}
\item Perl and XML::Parser
\item FFmpeg
\end{itemize}

The following command installs these on Debian:

\begin{verbatim*}
sudo apt-get install make \
  librsvg2-bin xsltproc \
  imagemagick gettext ffmpeg \
  git quilt zip \
  m4 automake \
  ttf-bitstream-vera fakeroot
\end{verbatim*}

\section{Target-specific Build Instructions}

\subsection{Compiling for Linux/UNIX}

The following additional packages are needed to build for Linux and
similar operating systems:

\begin{itemize}
\item \href{http://www.zlib.net/}{zlib}
\item \href{http://curl.haxx.se/}{CURL}
\item \href{http://www.lua.org/}{Lua}
\item \href{http://www.libsdl.org/}{SDL}
\item \href{http://www.libsdl.org/projects/SDL\_ttf/}{SDL\_ttf}
\item \href{http://www.libpng.org/}{libpng}
\item \href{http://libjpeg.sourceforge.net/}{libjpeg}
\item OpenGL (Mesa)
\item to run XCSoar, you need one of the following fonts (Debian
  package): DejaVu (\texttt{fonts-dejavu}),
  Roboto (\texttt{fonts-roboto}),
  Droid (\texttt{fonts-droid}),
  Freefont (\texttt{fonts-freefont-ttf})
\end{itemize}

The following command installs these on Debian:

\begin{verbatim*}
sudo apt-get install make g++ \
  zlib1g-dev \
  libfreetype6-dev \
  libpng-dev libjpeg-dev \
  libtiff5-dev libgeotiff-dev \
  libcurl4-openssl-dev \
  liblua5.2-dev lua5.2-dev \
  libxml-parser-perl \
  libasound2-dev \
  librsvg2-bin xsltproc \
  imagemagick gettext \
  mesa-common-dev libgl1-mesa-dev libegl1-mesa-dev \
  fonts-dejavu
\end{verbatim*}

To compile, run:

\begin{verbatim*}
make
\end{verbatim*}

You may specify one of the following targets with \texttt{TARGET=x}:

\begin{tabularx}{1.9\textwidth}{lX}

\texttt{UNIX} & regular build (the default setting) \\

\texttt{UNIX32} & generate 32 bit binary \\

\texttt{UNIX64} & generate 64 bit binary \\

\texttt{OPT} & alias for UNIX with optimisation and no debugging \\

\end{tabularx}

\subsection{Compiling for Android}

For Android, you need:

\begin{itemize}
\item \href{http://developer.android.com/sdk/}{Android SDK level 22}
\item \href{http://developer.android.com/sdk/ndk/}{Android NDK r19c}
\item \href{http://www.vorbis.com/}{Ogg Vorbis}
\item {Java JDK 
\begin{verbatim*}
sudo apt-get install default-jdk-headless vorbis-tools adb
\end{verbatim*}}
\end{itemize}

The required Android SDK components are:
\begin{itemize}
\item Android SDK Build-Tools 28.0.3
\item SDK Platform 22
\end{itemize}
These can be installed from the Android Studio SDK Manager, or using the SDK
command line tools:

\begin{verbatim*}
tools/bin/sdkmanager \
  "build-tools;28.0.3" \
  "platforms;android-22"
\end{verbatim*}

The \texttt{Makefile} assumes that the Android SDK is installed in
\verb|~/opt/android-sdk-linux| and the NDK is installed in
\verb|~/opt/android-ndk-r19c|.  You can use the options
\verb|ANDROID_SDK| and \verb|ANDROID_NDK| to override these paths.

Load/update the IOIO source code:

\begin{verbatim*}
git submodule init
git submodule update
\end{verbatim*}

To compile, run:

\begin{verbatim*}
make TARGET=ANDROID
\end{verbatim*}

Use one of the following targets:

\begin{tabularx}{1.9\textwidth}{lX}

\texttt{ANDROID} & for ARM CPUs (same as \texttt{ANDROID7}) \\

\texttt{ANDROID7} & for ARMv7 CPUs \\

\texttt{ANDROID7NEON} & with
\href{http://www.arm.com/products/processors/technologies/neon.php}{NEON}
extension \\

\texttt{ANDROID86} & for x86 CPUs \\

\texttt{ANDROIDMIPS} & for MIPS CPUs \\

\texttt{ANDROIDFAT} & "fat" package for all supported CPUs \\

\end{tabularx}

\subsection{Compiling for Windows}

To cross-compile to (desktop) Windows, you need
\href{http://mingw-w64.org}{Mingw-w64}.

The following command installs it on Debian:

\begin{verbatim*}
sudo apt-get install g++-mingw-w64
\end{verbatim*}

To compile for 32 bit Windows, run:

\begin{verbatim*}
make TARGET=PC
\end{verbatim*}

Use one of the following targets:

\begin{tabularx}{1.9\textwidth}{lX}

\texttt{PC} & 32 bit Windows (i686) \\

\texttt{WIN64} & Windows x64 (amd64 / x86-64) \\

\texttt{CYGWIN} & Windows build with Cygwin (experimental) \\

\end{tabularx}

\subsection{Compiling for iOS and macOS}

On macOS, the following tools are required:
\begin{itemize}
\item png2icns from \href{http://icns.sourceforge.net}{libicns} to build for
  macOS
\item \href{https://alioth.debian.org/projects/dpkg}{dpkg} to build the iOS
  IPA package
\item \href{http://cdrecord.org/private/cdrecord.html}{mkisofs} to build the
  macOS DMG package
\end{itemize}

To compile for iOS / AArch64, run:

\begin{verbatim*}
make TARGET=IOS64 ipa
\end{verbatim*}

To compile for iOS / ARMv7, run:

\begin{verbatim*}
make TARGET=IOS32 ipa
\end{verbatim*}

To compile for macOS / x86\_64, run:

\begin{verbatim*}
make TARGET=OSX64 dmg
\end{verbatim*}

\subsection{Compiling for the Raspberry Pi}

You need an ARM toolchain.  For example, you can use the Debian
package \verb|g++-arm-linux-gnueabihf|:

\begin{verbatim*}
make TARGET=PI
\end{verbatim*}

To optimize for the Raspberry Pi 2 (which has an ARMv7 with NEON
instead of an ARMv6):

\begin{verbatim*}
make TARGET=PI2
\end{verbatim*}

These targets are only used for cross-compiling on a (desktop)
computer.
If you compile on the Raspberry Pi, the default target will
auto-detect the Pi.

\subsection{Compiling for the Cubieboard}

To compile, run:

\begin{verbatim*}
make TARGET=CUBIE
\end{verbatim*}

This target is only used for cross-compiling on a (desktop) computer.
If you compile on the Cubieboard, the default target will auto-detect
the Cubieboard.

\subsection{Compiling for Kobo E-book Readers}

You need an ARM toolchain.  For example, you can use the Debian
package \verb|g++-arm-linux-gnueabihf|.

To compile XCSoar, run:

\begin{verbatim*}
make TARGET=KOBO
\end{verbatim*}

To build the kobo install file \texttt{KoboRoot.tgz}, you need the
following Debian packages:

\begin{verbatim*}
sudo apt-get install fakeroot ttf-bitstream-vera
\end{verbatim*}

Then compile using this command:

\begin{verbatim*}
make TARGET=KOBO output/KOBO/KoboRoot.tgz
\end{verbatim*}

For this, you need the Debian package \verb|libc6-armhf-cross|.

\subsubsection{Building USB-OTG Kobo Kernel}

To build a USB-OTG capable kernel for the Kobo, clone the git
repository:

\begin{verbatim*}
git clone git://git.xcsoar.org/xcsoar/max/linux.git
\end{verbatim*}

Check out the correct branch. For the Kobo Mini, this is the "kobo" branch,
for the Kobo Glo HD, the branch is called "kobo-glohd", and for the Kobo Aura 2,
use the branch "kobo-aura2".

\begin{verbatim*}
git checkout kobo
\end{verbatim*}

Configure the kernel using the configuration files from the \texttt{kobo/kernel}
directory in XCSoar's \texttt{git} repository. For the Kobo Mini, install a
\href{http://openlinux.amlogic.com:8000/download/ARM/gnutools/arm-2010q1-202-arm-none-linux-gnueabi-i686-pc-linux-gnu.tar.bz2}{gcc
  4.4 cross compiler}, for example in \texttt{/opt}. For the Kobo Glo HD and
Aura 2, install a
\href{https://launchpad.net/gcc-arm-embedded/4.6/4.6-2012-q4-update/+download/gcc-arm-none-eabi-4_6-2012q4-20121016.tar.bz2}{gcc
  4.6 cross compiler}

To compile a kernel image for the Kobo Mini, type:

\begin{verbatim*}
make \
  CROSS_COMPILE=/opt/arm-2010q1/bin/arm-none-linux-gnueabi- \
  ARCH=arm uImage
\end{verbatim*}

To compile a kernel image for the Kobo Glo HD, type:

\begin{verbatim*}
make \
  CROSS_COMPILE=/opt/gcc-arm-none-eabi-4_6-2012q4/bin/arm-none-eabi- \
  ARCH=arm uImage
\end{verbatim*}

Copy \texttt{uImage} to the Kobo.  Kernel images can be installed with
the following command:

\begin{verbatim*}
dd if=/path/to/uImage of=/dev/mmcblk0 bs=512 seek=2048
\end{verbatim*}

Note that XCSoar's \texttt{rcS} script may overwrite the kernel image
automatically under certain conditions.  To use a new kernel
permanently, install it in \texttt{/opt/xcsoar/lib/kernel}.  Read the
file \texttt{kobo/rcS} to find out more about this.

To include kernel images in \texttt{KoboRoot.tgz}, copy
\texttt{uImage.otg}, \texttt{uImage.kobo}, \texttt{uImage.glohd.otg},
\texttt{uImage.glohd}, \texttt{uImage.aura2} and \texttt{uImage.aura2.otg} to
\texttt{/opt/kobo/kernel}.

\subsection{Editing the Manuals}

The XCSoar documententation, including the Developer Manual that you are 
reading right now, is written using the TeX markup language. You can edit
the source files with any text editor, although a specific TeX editor (e.g. LateXila) 
makes it easier. 

Source files are located in the en, fr, de, pl subdirectories of the doc/manual directory. 
The Developer manual is in the doc/manual/en directory. The generated files are put into
the output/manual directory.

To generate the PDF manuals, you need the TexLive package, plus some European languages.

The following command installs these on Debian:

\begin{verbatim*}
sudo apt-get install texlive \
  texlive-latex-extra \
  texlive-luatex \
  texlive-lang-french \
  texlive-lang-polish \
  texlive-lang-german \
  texlive-lang-portuguese \
  liblocale-po-perl
\end{verbatim*}

The documentation is distributed as PDF files. Generating the PDFs from the 
TeX files is done by typing:

\begin{verbatim*}
make manual
\end{verbatim*}

A lot of warnings are generated... this is normal. Check for the presence of PDF files 
to ensure that the generation process was successful.
\section{Options}

\subsection{Parallel Build}

Most contemporary computers have multiple CPU cores.  To take
advantage of these, use the \texttt{make -j} option:

\begin{verbatim*}
make -j12
\end{verbatim*}

This command launches 12 compiler processes at the same time.

Rule of thumb: choose a number that is slightly larger than the number
of CPU cores in your computer.  12 is a good choice for a computer
with 8 CPU cores.

\subsection{Optimised Build}

By default, debugging is enabled and compiler optimisations are
disabled.  The resulting binaries are very slow.  During development,
that is helpful, because it catches more bugs.

To produce optimised binaries, use the option \texttt{DEBUG}:

\begin{verbatim*}
make DEBUG=n
\end{verbatim*}

Be sure to clean the output directory before you change the
\texttt{DEBUG} setting, because debug and non-debug output files are
not compatible.

The convenience target \texttt{OPT} is a shortcut for:

\begin{verbatim*}
TARGET=UNIX DEBUG=n TARGET_OUTPUT_DIR=output/OPT
\end{verbatim*}

It allows building both debug and non-debug incrementally, because two
different output directories are used.

\subsection{Compiling with ccache}

To speed up the compilation of XCSoar we can use \texttt{ccache} to cache the 
object files for us. All we have to do is install ccache and 
add 
\texttt{USE\_CCACHE=y} 
to the make command line:

\begin{verbatim*}
sudo apt-get install ccache
make TARGET=UNIX USE_CCACHE=y
\end{verbatim*}

\section{Using a build VM with Vagrant}\label{sec:vagrant}
An easy way to install a virtual machine with all build dependencies required
for various targets (e. g. Linux, Windows, Android and Kobo), is using Vagrant.

The following is needed to install the VM with Vagrant:

\begin{itemize}
\item \href{https://www.vagrantup.com/}{Vagrant}
\item \href{https://www.virtualbox.org/}{VirtualBox}
\end{itemize}

The Vagrantfile can be found in the \texttt{ide/vagrant} subfolder of the \xc
source. To set up the VM, and connect to it, type:

\begin{verbatim*}
cd ide/vagrant
vagrant up
vagrant ssh
\end{verbatim*}

The XCSoar source directory on the host is automatically mounted as a shared
folder at \texttt{/xcsoar-host-src} in the VM. For performance reasons, it is
not recommended to compile directly in this folder. A git clone of this
directory is automatically created in the home directory
(\texttt{~/xcsoar-src}), which should be used instead. In this git clone, the
XSoar source directory on the host is preconfigured as a git remote named
"host", and the XCSoar master directory is preconfigured as a remote named
"master".

To shutdown the VM, type:

\begin{verbatim*}
vagrant halt
\end{verbatim*}
