\documentclass[a4paper,12pt]{refrep}
\input{xcsoar.sty}

\title{Developer Manual}

\input{xcsoar-headers.sty}
\xcsoarheader{XCSoar Developer Manual}

\input{title-xcsoar.sty}
\usepackage{listings}

\usepackage{tikz}
\usetikzlibrary{arrows,shapes,fit,decorations,decorations.pathmorphing,decorations.pathreplacing,calc}

\tikzstyle{thread}=[draw, fill=lightgray, text centered,
  minimum width=7em, minimum height=2.5em]

\begin{document}
\maketitle

\input{toc.tex}

%%%%%%%%%%%%%%%%%%%%%%

\chapter*{Preface}

This manual applies to XCSoar version 6.0.  The authors reserve the
right to update this manual as enhancements are made throughout the
life of this product.

\section*{Warnings and precautions}

\warning IT IS THE USER'S RESPONSIBILITY TO USE THIS SOFTWARE PRUDENTLY. THIS
SOFTWARE IS INTENDED TO BE USED ONLY AS A NAVIGATION AID AND MUST NOT
BE USED FOR ANY PURPOSE REQUIRING PRECISE MEASUREMENT OF DIRECTION,
DISTANCE, LOCATION, OR TOPOGRAPHY. THIS SOFTWARE SHOULD NOT BE USED AS
AN AID TO DETERMINE GROUND PROXIMITY FOR AIRCRAFT NAVIGATION.
THIS SOFTWARE SHOULD NOT BE USED AS A TRAFFIC COLLISION AVOIDANCE SYSTEM.


\section*{Legal notices}

\subsection*{Software license agreement}

This software is released according to the GNU General Public License
Version~2.  See Appendix~\ref{cha:gnu-general-public} for the full
text of the agreement and warranty notice.

\subsection*{Limited liability}

In no event shall XCSoar, or its principals, shareholders, officers,
employees, affiliates, contractors, subsidiaries, or parent
organizations, be liable for any incidental, consequential, or
punitive damages whatsoever relating to the use of the Product.

\subsection*{Disclaimer}

This product, and all accompanying files, data and materials, are
distributed "as is" and with no warranties of any kind, whether
express or implied.  This product is used entirely at the risk of the
user.  Although great care has been taken to eliminate defects during
its development it is not claimed to be fault-free. No claims are made
regarding its correctness, reliability or fitness for any particular
purpose.  The XCSoar project developers and contributors shall not be
liable for errors contained herein or for incidental or consequential
damages, loss of data or personal injury in connection with
furnishing, performance, or use of this material.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}\label{cha:introduction}

\input{compiling.tex}

\chapter{Policy}\label{cha:policy}

\section{Writing Patches}

Submit patches or \texttt{git pull} requests to the developer mailing
list (\texttt{xcsoar-devel@lists.sourceforge.net}).  We provide
accounts on \texttt{git.xcsoar.org} to regular contributors.

A patch should be self-explanatory, it needs a good description.  The
subject line specifies the subsystem/library name and a brief
description of what is changed, followed by an empty line.  Then write
a longer description if needed, and explain why this change is needed.

Each patch must compile and must not introduce a regression (as far as
we know at the time).

Each patch must be self-contained and should only change one thing.
Split larger patches into smaller pieces.  Don't refactor and
add/modify/remove features in the same patch.

Don't rewrite code unless you need to.  Migrate incrementally to a new
concept.  Keep patches small and easy to understand.

Patch files can be generated by running
\begin{verbatim*}
git diff > patch
\end{verbatim*}

\section{Code Style}

79 columns, reasonable exceptions allowed.  Indent 2 spaces, no tabs.
No indent for \verb|namespace| blocks (a compromise to avoid excessive
indentation).

\emph{Comments:} write enough code comments (in English).  All
workarounds must be documented.  Everybody must be able to understand
your code, even when you're gone.  Don't abuse multiple single-line
comments (``//'') to write mult-line comments.

\emph{API documentation:} non-trivial functions should be documented
in a doxygen comment.

\emph{Names:} class/function names in \texttt{CamelCase} (not
\texttt{camelCase}); attributes/variables lower case, separated with
underscore (e.g. \texttt{foo\_bar}); constants (including
\texttt{enum} values) all upper case (e.g. \texttt{FOO\_BAR}).

Exception: when a foreign API is being mimicked (e.g. STL containers),
we adopt its naming conventions.

\emph{Files:} \texttt{*.cpp} and \texttt{*.hpp} for C++.  Files should
be named after the main class which is provided.  Each class should
have a separate source file and a separate header.  UNIX text format.

Be \texttt{const}-correct.  Use \verb|constexpr| instead of
\verb|const| whenever possible.

Use \verb|static| whenever possible.  Functions and global variables
that are only used in one source file should not be exported.  Methods
that do not use any instance method/variable should be static to avoid
the overhead of passing the implicit \verb|this| parameter.

Make methods \verb|virtual| only after careful consideration.  A
destructor should only be virtual if necessary.  All overrides must
use the \verb|override| keyword.  Use \verb|final| often.

Compile with \texttt{WERROR=y} and fix all warnings.

Don't write large functions.  Split them up when they become too
large.

Avoid dynamic allocation.  Dynamic allocation means overhead, more
locking and heap fragmentation.  Use \texttt{StaticArray} and
\texttt{StaticString} if possible.

Asterisks belong to the variable name, not to the type name.  Consider
``\texttt{Foo* a, b}''.  ``\texttt{Foo *a, b}'' or ``\texttt{Foo *a,
  *b}'' is easier to understand.

Some sample code to demonstrate our code style:

\begin{verbatim}
/**
 * API documentation for this class.
 */
struct TheStruct {
  unsigned an_attribute;
  bool second_attribute;

  TheStruct();

  /**
   * API documentation for this method.
   *
   * @param foo documentation for this parameter
   * @return documentation for the return value
   */
  bool TheMethod(int foo);
};

TheStruct::TheStruct()
  :an_attribute(0),
   second_attribute(true)
{
}

static bool
FooBar(int a_parameter, unsigned another_parameter,
       const TheStruct *next_row)
{
  switch (a_parameter) {
  case 0:
    break;
  }

  if (a_parameter == 2 && another_parameter == 3 &&
      next_row != NULL)
    return true;

  return a_parameter == 42;
}
\end{verbatim}

\section{C++}

XCSoar is written in C++14.

XCSoar's standard compilers are \texttt{gcc} (at least version 4.8.1) and
\texttt{clang} (at least version 3.4).

\subsection{Other rules}

In a class declaration, attributes come first, then
constructor/destructor, and finally the methods.  Having all
attributes in one place gives a good overview of the nature of a
class.

Avoid expensive and bloated STL containers if there are cheaper
solutions (e.g. \texttt{StaticArray}, \texttt{StaticString} if the
maximum size is predictable).

Avoid template hell.  Keep templates readable.  Keep in mind that
excessive template use may bloat the binary.

\section{Graphical User Interface}
\subsection{Letter Cases}

Following the guidline should prevent the GUI from mixtures of "ON" and 
"On" text elements, and lead to a systematic GUI text presentation. 
The goal is to recognize GUI text fast and reliable.

\begin{description}
  \item[Captions] Captions (button captions, windows titles) to use
        capitalization. E.g. ,"Pan On", "The Display Of ...". 

  \item[Abbreviations] Generally known abbreviation use upper case like "MC", "ETA",
        "V"; or they can use CamelCase, especially when using synthetic
        words  like "GoTo", "InfoBox". 
        Abbreviated words by simply cutting the end of the word needs a
        dot, e.g. "Max. temp."

  \item[Plain text] Longer help texts are to write like prose: "This is the help
        page for ...".

  \item[Labels] Label text has the least systematic constraints:
\begin{itemize}
    \item Captions for text (input) fields, e.g. "Wing loading"
    \item Info text on widgets. E.g. "No data" on an empty
          analysis page. 
    \item Label text for radio or check boxes. 
    \item Selections on Combo-boxes, selectors, Pull-down menus. 
\end{itemize}
  All those should go like prose, whereas exceptions might be
  meaningful.
  
  \item[Gauge caption] Also the appearance of the gauge caption should be covered with
        that. They are currently mapped to upper case all over. I think
        the most readable also here is a CamelCase approach. E.g. to
        distinct "WP Dist", "WP AltD", and "WP AltR". Another good
        example would be MACCREADY, which should be MacCready, or just
        MC.

  \item[Units] Units have their own specific appearance. A profound paper is
        http://physics.nist.gov/cuu/pdf/checklist.pdf we could just
        refer to.
\end{description}


\chapter{Architecture}

This chapter describes XCSoar's internal code architecture.

\section{Source Organisation}

XCSoar's source code is stored in the \texttt{src} directory.  This
section tries to give a rough overview where you can find what.

\begin{itemize}

\item \texttt{Util/}: generic C++ utilities that do not depend on
  external libraries, such as data structures, string operations

\item \texttt{Math/}: math data types (fixed-point math, angles) and
  generic formulas

\item \texttt{Geo/}: geographic data structures and formulas

\item \texttt{Formatter/}: code that formats internal values to
  strings

\item \texttt{Units/}: conversion from SI units (``System'' units) to
  configured user units

\item \texttt{NMEA/}: data structures for values parsed from NMEA

\item \texttt{Profile/}: user profiles, loading from and saving to

\item \texttt{IGC/}: support for the IGC file format

\item \texttt{Logger/}: all loggers (NMEA, IGC, flights)

\item \texttt{Thread/}: multi-threading support (OS specific)

\item \texttt{Screen/}: base library for the graphical user interface

\item \texttt{Renderer/}: various graphical renderers, for map and
  analysis

\item \texttt{MapWindow/}: the map

\item \texttt{Form/}: modal dialogs and their controls (based on the
  screen library)

\item \texttt{Dialogs/}: modal dialogs implementations (based on the
  form library)

\item \texttt{Net/}: networking code (OS specific)

\item \texttt{Operation/}: generic code to support cancellable
  long-running operations

\item \texttt{Android/}: code specific to Android (the native part
  only; Java code is in \texttt{android/src/}

\item \texttt{Engine/PathSolvers/}: an implementation of Dijkstra's
  path finding algorithm, for task and contest optimisation

\item \texttt{Engine/Airspace/}: airspace data structures and airspace
  warnings

\item \texttt{Engine/Waypoint/}: waypoint data structures

\item \texttt{Engine/GlideSolvers/}: a MacCready implementation

\item \texttt{Engine/Task/}: task data structures and calculations

\item \texttt{Engine/Contest/}: contest optimisation

\item \texttt{Engine/Route/}: the route planner (airspace and terrain)

\end{itemize}

\section{Threads and Locking}

\subsection{Threads}

XCSoar runs on multiple threads, to make the UI responsive but still
allow expensive background calculations.

This is how it looks like on Windows and Linux/SDL (software
rendering):

\begin{tikzpicture}
  \node(ui)[thread]{UI thread};

  \path(ui.east)+(4,0) node(draw)[thread]{DrawThread};
  \draw[->] (ui.10) -- (draw.170)
  node[midway, above, sloped, font=\tiny]{redraw (Pan)};
  \draw[->] (draw.190) -- (ui.350)
  node[midway, below, sloped, font=\tiny]{BufferCanvas};

  \path(draw.south)+(0,-2) node(calc)[thread]{CalcThread};
  \draw[->] (calc.north) -- (draw.south)
  node[midway, below, sloped, font=\tiny]{results};

  \path(calc.south)+(0,-2) node(merge)[thread]{MergeThread};
  \draw[->] (merge.north) -- (calc.south)
  node[midway, below, sloped, font=\tiny]{results};

  \path(merge.south)+(-3,-2) node(device1)[thread]{Device};
  \draw[->] (device1.north) -- (merge.250)
  node[midway, above, sloped, font=\tiny]{sensor data};

  \path(merge.south)+(3,-2) node(device2)[thread]{Device 2};
  \draw[->] (device2.north) -- (merge.290)
  node[midway, above, sloped, font=\tiny]{sensor data};

  \path(ui.south)+(0,-3) node(io)[thread]{I/O thread};
  \draw[->] (io.north) -- (ui.south)
  node[midway, above, sloped, font=\tiny]{data};
  \draw[->] (io.east) -- (merge.west)
  node[midway, above, sloped, font=\tiny]{sensor data};
\end{tikzpicture}

The UI thread is the main thread.  It starts the other threads and is
responsible for the UI event loop.  No other thread is allowed to
manipulate windows.  The UI thread has a timer which does regular
house keeping twice per second (\texttt{Pro\-cess\-Ti\-mer.cpp}).

The calculation thread (\texttt{Calcu\-la\-tion\-Thread.cpp},
\texttt{Glide\-Com\-pu\-ter*.cpp}) does all the expensive calculations
in background.  It gets data from the devices (through
\texttt{Merge\-Thread}) and forwards it together with calculation
results to the drawing thread and the main thread.

Each device has its own thread (\texttt{Serial\-Port.cpp}).  This is
needed because Windows CE does not support asynchronous COMM port I/O.
The thread is stopped during task declaration (which happens in the UI
thread).

When new data arrives on the serial port, the \texttt{Merge\-Thread}
gets notified, which will merge all sensor values into one data
structure.  It will then run cheap calculations, and forwards
everything to the \texttt{Calcu\-la\-tion\-Thread}.

With OpenGL, the map is rendered live without a buffer.  There is no
DrawThread.

On Android, the UI thread is not the main thread - the main thread is
implemented in Java, managed by Android itself.  The UI thread listens
for events which the Java part drops into the event queue
(\texttt{NativeView.java} and others).  The internal GPS does not need
a thread, it is implemented with Java callbacks.  For Bluetooth I/O,
there are two threads implemented in Java (\texttt{InputThread.java}
and \texttt{OutputThread.java}, managed by
\texttt{BluetoothHelper.java}).

\subsection{Locking}

Some data structures are rarely modified.  There is no lock for them.
For a modifications, all threads must be suspended.  Example:
waypoints, airspaces.

Other data structures are modified so often that correct locking would
be too much overhead.  Each thread and each instance has its own
copy.  The lock needs to be obtained only for making the private
copy.  The private copy can be used without locking.  Example:
\texttt{NMEA\_INFO}, \texttt{DERIVED\_INFO}.

There are objects which are too expensive to copy.  Normal locking
applies to them.  We have a template class called \texttt{Guard} to
enforce proper read/write locking.  Example: the task.


\section{Accessing Sensor Data}

Much of XCSoar deals with obtaining sensor data and visualising it.

Suppose you want to write a dialog that needs the current GPS
location, where do you get it?  The short and simple answer is: from
\verb|CommonInterface::Basic()| (the \texttt{InterfaceBlackboard}).
Example:

\begin{verbatim}
#include "Interface.hpp"

...
  const auto &basic = CommonInterface::Basic();
  if (basic.location_available)
    current_location = basic.location;
\end{verbatim}

This is true for the main thread (aka the ``user interface thread'').
Other threads must not use the \texttt{Interface.hpp} library, because
the \verb|InterfaceBlackboard| is not protected in any way.  It
contains copies of various data structures just for the main thread.

This is how sensor data moves inside XCSoar:

\begin{tikzpicture}
  \node(merge)[thread,align=center]{
    MergeThread\\
    BasicComputer\\
    DeviceBlackboard
  };

  \path(merge.north)+(-2,2) node(device1)[thread]{Device 1};
  \draw[->] (device1.south) -- (merge.north)
  node[midway, below, sloped, font=\tiny]{NMEAInfo};

  \path(merge.north)+(2,2) node(device2)[thread]{Device 2};
  \draw[->] (device2.south) -- (merge.north)
  node[midway, below, sloped, font=\tiny]{NMEAInfo};

  \path(merge.south)+(0,-2) node(calc)[thread,align=center]{
    CalcThread\\
    GlideComputer\\
    GlideComputerBlackboard
  };
  \draw[->] (merge.south) -- (calc.north)
  node[midway, below, sloped, font=\tiny]{MoreData};

  \path(calc.south)+(0,-2) node(ui)[thread,align=center]{
    UI thread\\
    InterfaceBlackboard\\
    BlackboardListener
  };
  \draw[->] (calc.south) -- (ui.north)
  node[midway, below, sloped, font=\tiny]{DerivedInfo};

  \path(ui.east)+(3.5,0) node(draw)[thread,align=center]{
    DrawThread\\
    MapWindow\\
    MapWindowBlackboard
  };
  \draw[->] (ui.east) -- (draw.west)
  node[midway, below, sloped, font=\tiny]{DerivedInfo};
\end{tikzpicture}

The device driver parses input received from its device into its own
\texttt{NMEAInfo} instance inside \texttt{DeviceBlackboard}
(i.e. \verb|per_device_data|).  Then it wakes up the
\texttt{MergeThread} to merge the new data into the central
\texttt{NMEAInfo} instance.  The \texttt{MergeThread} hosts the
\texttt{BasicComputer} which attempts to calculate missing data (for
example, derives vario from GPS altitude).

The \texttt{CalculationThread} wakes up and receives the
\texttt{MoreData} object from \texttt{DeviceBlackboard}.  Here,
expensive calculations are performed (\texttt{GlideComputer}: task
engine, airspace warnings, ...), resulting in a \texttt{DerivedInfo}
object.  The \texttt{CalculationThread} runs no more than twice per
second.

Finally, the UI thread wakes up and receives \texttt{MoreData} and
\texttt{DerivedInfo} via \texttt{DeviceBlackboard}.  This updates
InfoBoxes and other UI elements.  On Windows, the map is drawn in a
separate thread, so there's another layer.

Let's get back to the question: where do I get sensor data?  That
depends on who you are:

\begin{itemize}

\item you are the user interface: (InfoBoxes, dialogs, any Window
  callback): \texttt{InterfaceBlackboard} (see above).  To get
  notified on changes, register a \texttt{BlackboardListener} (and
  don't forget to unregister it).

\item you are the MapWindow: depends!  If you're being called from
  \texttt{OnPaintBuffer} (i.e. inside the \texttt{DrawThread}), you
  must use the \texttt{MapWindowBlackboard}, all others must use the
  \texttt{InterfaceBlackboard}.

\item you are a ``computer'' library: you will get the values as a
  parameter.  Don't try to use the \texttt{GlideComputerBlackboard}
  directly.

\item you are a device driver: implement the method
  \texttt{OnSensorUpdate} or \texttt{OnCalculatedUpdate} if you need
  to know values from other devices or calculation results.

\item everybody else may use the \texttt{DeviceBlackboard}, but be
  sure to lock it while using its data.

\end{itemize}


\chapter{The build system}\label{cha:build}

A big plain \texttt{Makefile} is used to control the XCSoar build.
GNU extensions are allowed.

This chapter describes the internals of our build system; for
instructions on compiling XCSoar, see chapter \ref{cha:compiling}.

\section{Linker parameters}

The following variables (or variable suffixes) appear in the
\texttt{Makefile} (conforming to \texttt{automake} conventions):

\begin{description}
\item[\texttt{LDFLAGS}] Linker flags, such as \texttt{-static} or
  \texttt{-Wl,...}, but not \texttt{-l}.
\item[\texttt{LDLIBS}] All \texttt{-l} flags, e.g. \texttt{-lGL}.
\item[\texttt{LDADD}] Path names of static libraries,
  e.g. \texttt{/usr/lib/libz.a}.
\end{description}

Search directories (\texttt{-L}) are technically linker ``flags'', but
they are allowed in \texttt{LDLIBS}, too.


\chapter{Developing}

\section{Debugging XCSoar}

The XCSoar source repository contains a module for the GNU debugger
(\texttt{gdb}).  It contains pretty-printers for various XCSoar types,
including \texttt{Angle}, \texttt{GeoPoint} and others.  These are
helpful when you print values in the debugger.  To use it, start the
debugging session and load the module:

\begin{verbatim}
$ gdb -ex "source tools/gdb.py" output/UNIX/xcsoar
(gdb) run
\end{verbatim}

The module will automatically convert fixed-point to floating point,
radian angles to degrees and more.  You can now do fancy stuff like:

\begin{verbatim}
(gdb) p basic.location
$1 = GeoPoint(7.93911242887 51.1470221074)
(gdb) p basic.date_time_utc
$2 = DateTime(2012/12/23 21:41:57)
(gdb) p basic.track
$3 = 55.2254197961
(gdb) p basic.external_wind
$4 = GeoVector::ZERO
(gdb) p current_leg.vector_remaining
$5 = GeoVector(267.899420345 107957.109724)
\end{verbatim}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{User interface guidelines}

\section{General}

\begin{itemize}
\item Minimise the number of colours, and re-use colour groups already defined.
\item Too much use of colour where it is not required serves only to reduce 
 the effectiveness of bright colours for important items.
\item High colour saturation elements should be reserved for high importance items
\item High contrast against background should be reserved for high importance items
\item Attempt to adopt colours that are intuitive based the function of the item
\item Minimise the clutter where possible --- readibility is essential for use 
 in flight
\item Use colours defined in \verb|Graphics| according to functional name, not 
 their actual colour.
\item Try to maintain consistent use of colours in all uses of that function, 
 such as dialogue graphics as well as map overlays and infoboxes.
\item Text should always be monochrome.
\end{itemize}

Use aviation conventions or adopt best aviation human factors
standards where possible, in particular:
\begin{itemize}
\item ICAO Internation Standards and Recommended Practices, Annex 4 to the 
 Convention on International Civil Aviation (Aeronautical Charts).
\item NASA Colour Usage recommendations and design guidelines: \verb|http://colorusage.arc.nasa.gov/|
\item DOT/FAA/AR-03/67 {\em Human Factors Considerations in the Design and 
 Evaluation of Electronic Flight Bags (EFBs)} \verb|http://www.volpe.dot.gov/hf/aviation/efb/docs/efb_version2.pdf|
\item FAA Human Factors Design Standards \verb|http://hf.tc.faa.gov/hfds/|.
\item DOT/FAA/AM-01/17 {\em Human Factors Design Guidelines for Multifunction Displays}
\end{itemize}

Check for performance with respect to colour blindness.
This site has a useful tool that can be used to convert screenshots
to how they would look to a person with common color blindness:
\verb|http://www.etre.com/tools/colourcheck/|.

{\bf For safety purposes, avoid use of elements that may encourage or require 
the user to stare at the screen continuously.}

{\bf For safety purposes, avoid user controls that have significant risk of 
producing unsafe results if misconfigured by the pilot.}

\subsection{General colour conventions}
Colour conventions generally in use throughout the program:
\begin{itemize}
\item Red for indicator of warning
\item Orange for indicator of caution
\item Green for positive indicator of safety
\item Blue for neutral indicator of safety
\end{itemize}

\subsection{Displayed data}
\begin{itemize}
\item Where data is invalid, indicate this by not presenting the data or
  showing dashes.
\item Present data in user-defined units.
\item Display numerical data with significant digits appropriate to the accuracy of the
  calculations, or its functional use by the pilot, whichever is lower.
\end{itemize}

\section{Dialogs and menu buttons}

\subsection{Colors}
Colour conventions in use are:
\begin{itemize}
\item Grey for buttons
\item Buttons and other widgets rendered with an evenly shaded border
\item Yellow for clicked items
\item Light blue for the key focused item
\item Medium blue for dialogue title bar
\item Text is black if the item is enabled
\item Text is greyed out (but still visible) if the item is disabled
\end{itemize}

\subsection{dialogue types and navigation buttons}
There are four types of dialogs in XCSoar, and the navigation 
buttons for each are different.  
Navigation buttons are the Close, OK, Cancel and Select buttons.
\begin{itemize}
\item Dialogs that modify and save data when the dialogue closes.

These shall usually have a Close button (no Cancel) and may have context specific 
function buttons
\item Dialogs that modify data where Cancel would be important for the user.

These shall have OK and Cancel buttons.  This may include dialogs with 
children dialogs where hitting Cancel from the parent dialogue cancels 
all the changes made in the children dialogs

\item Dialogs that have a list of values, one of which can be selected to 
 return to the parent dialogue.

These shall have Select and Cancel buttons
\item Dialogs that display information that cannot be modified.

These shall have a Close button
\end{itemize}

\subsection{dialogue button placement and size}
\begin{itemize}
\item The Close and Cancel buttons will never appear in the same dialogue and 
 are always located in the same place.  This location will be:

For portrait: lower right

For landscape: lower left
\item The Select button will be accompanied with a Cancel button.  The locations will be:

For portrait: Select in lower left, Cancel in lower right

For landscape: Cancel in lower left, Select immediately above it
\item Buttons will be 35 (scaled) pixels high
\item Buttons will be flush with the bottom of the screen and with the sides of 
 the screen and against each other (no margins)
\item In portrait, buttons will be 33% or 50% or 100% width of the screen
\item In landscape, buttons will be 65 to 80 (scaled) pixels wide, as wide as the 
 frame permits.  They will generally be a vertical row of buttons flush left of the screen
\item If text won't fit on a button, the buttons can be made larger consistently 
 for a screen, but this should be the exception because if it must contain that 
 much text consider using a different type of control.
\item Exceptions to all the dialogue concepts above are encouraged, but should be 
 mocked up and reviewed with the development community prior to implementing and 
 possibly documenting in the developers guide.
\end{itemize}

\subsection{Usability}
\begin{itemize}
\item Minimum size of buttons should be X by Y mm
\item Ensure all dialogs are navigable using cursor keys only
\item Ensure the focussed item is clearly identified.  The rectangle
  of the widget on the canvas may be drawn using the \verb|fill_focus| method
  of \verb|Canvas|.
\end{itemize}

\section{Main graphics}

\subsection{Colors}
Colour conventions in use, in order of priority, are:
\begin{itemize}
\item Aircraft black and white, for neutrality but clear identification
\item Traffic (FLARM) use alarm green, orange, and red.
\item Lift is vibrant green, sink is copper orange.
\item Aircraft navigation (route, best cruise track) is (ICAO) dark purple-blue
\item Task navigation lines and areas are (ICAO) magenta.
\item Updraft sources and other updraft derived data is sky blue.
\end{itemize}

(Todo) airspace alert colours

Map culture (topography) and terrain rendering should conform to ICAO
Annex 4 where appropriate.  Note that some modifications are
reasonable for electronic use given that Annex 4 deals with paper
charts.  Nevertheless, the colour conventions are useful to adopt as they are
likely to be intuitive and are designed for aviation use.

\subsection{Pen styles}

\begin{itemize}
\item Map culture should be rendered with a thin pen
\item Thicker pens used for important (e.g.\ task, navigational, airspace) lines
\item Dashed lines are used to increase perceptual priority
\end{itemize}

\subsection{Map overlays}
Elements on the map that are not part of the map layer, such as additional
informational widgets (final glide bar, wind, north arrow) should be rendered
so as to help those elements be visually separated from the map:

\begin{itemize}
\item Generally adopt higher contrast (higher colour saturation or darker shade) 
 than the background map layer elements.

\item For elements covering an area (non line), draw the entire element or a border
with a luminosity contrasting pen, of width \verb|IBLSCALE(1)|.

\item Consider whether the widget is required in all flying states and display modes.
if it does not serve a direct functional purpose in some states/modes, do not
render it.

\item Avoid locating widgets at the aircraft symbol (ownship symbol).
It is important to keep this area clear so the aircraft symbol can be easily found.

\end{itemize}

Elements that may be rendered over each other should be organised in order of
priority, particularly with alert warning items above caution items above non-alert items.

\section{Terminology}
\subsection{Glide Ratio}
'Glide ratio' is a non-specific term which can refer to the ratio of horizontal to 
vertical motion with reference to either the surrounding airmass or the ground.

To reduce confusion, ground-referenced glide ratios (eg distance travelled over 
ground vs altitude lost) should be referred to by the term 'glide ratio over 
ground' when space allows, or 'glide ratio' / 'GR'.

Air-referenced glide ratios (eg airspeed vs sink rate) should be specified as 
'lift/drag ratio' / 'L/D ratio' / 'LD'. The lift/drag ratio is numerically equal 
to the air-referenced glide ratio when flying at constant speed.

If usage spans both air-referenced and ground-referenced glide ratios, the 
non-specific term 'glide ratio' / 'GR' should be used. 'Lift/drag ratio' should 
never be used to refer to ground-referenced glide ratios.

\chapter{Lua Scripting}\label{cha:lua}
\input{lua.tex}

\chapter{File formats}\label{cha:file_formats}

\input{input_events.tex}
\input{tpl_format.tex}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix

\chapter{Setting up a development environment based on linux}\label{cha:developmentsetup}
\input{developsetuplinux.tex}

\chapter{GNU General Public License}\label{cha:gnu-general-public}
\input{gpl.tex}

\end{document}
