\documentclass[a4paper,11pt]{refrep}
\input{xcsoar.sty}

% Set the title
\title{Podręcznik}

% Set the page title
\input{xcsoar-headers.sty}
\xcsoarheader{Podręcznik dla XCSoar}

% Define some xc doc styles
\input{title-xcsoar.sty}

% Select the right encoding and language package
\usepackage[polish]{babel}
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%
% Front page
\maketitle

%%%%%%%%%%%%%%%%%%%%%%
% Todo's
\listoftodos
 
%%%%%%%%%%%%%%%%%%%%%%
% Table of contents
\input{toc.tex}

%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Preface}

\section*{Warnings and precautions}

\warning IT IS THE USER'S RESPONSIBILITY TO USE THIS SOFT\-WARE PRUDENTLY. THIS SOFTWARE IS 
INTENDED TO BE USED ONLY AS A NAVIGATION AID AND MUST NOT BE USED FOR ANY PURPOSE REQUIRING 
PRECISE MEASURE\-MENT OF DIRECTION, DISTANCE, LOCATION, OR TOPO\-GRAPHY. THIS SOFTWARE SHOULD 
NOT BE USED AS AN AID TO DETERMINE GROUND PROXIMITY FOR AIRCRAFT NAVIGATION. THIS SOFTWARE 
SHOULD NOT BE USED AS A TRAFFIC COLLISION AVOIDANCE SYSTEM.


\section*{Legal notices}

\subsection*{Software license agreement}

This software is released according to the GNU General Public License
Version~2.  See Appendix~\ref{cha:gnu-general-public} for the full
text of the agreement and warranty notice.

\subsection*{Limited liability}

In no event shall XCSoar, or its principals, shareholders, officers,
employees, affiliates, contractors, subsidiaries, or parent
organizations, be liable for any incidental, consequential, or
punitive damages whatsoever relating to the use of the Product.

\subsection*{Disclaimer}

This product, and all accompanying files, data and materials, are
distributed "as is" and with no warranties of any kind, whether
express or implied.  This product is used entirely at the risk of the
user.  Although great care has been taken to eliminate defects during
its development it is not claimed to be fault-free. No claims are made
regarding its correctness, reliability or fitness for any particular
purpose.  The XCSoar project developers and contributors shall not be
liable for errors contained herein or for incidental or consequential
damages, loss of data or personal injury in connection with
furnishing, performance, or use of this material.


%%%%%%%%%%%%%%%%%%%%%%
\input{ch01_introduction.tex}
\input{installation.tex}

%%%%%%%%%%%%%%%%%%%%%%
\input{ch02_user_interface.tex}

%%%%%%%%%%%%%%%%%%%%%%
\input{ch03_navigation.tex}

%%%%%%%%%%%%%%%%%%%%%%
\input{ch04_xc_tasks.tex}

%%%%%%%%%%%%%%%%%%%%%%
\input{ch05_glide_computer.tex}

%%%%%%%%%%%%%%%%%%%%%%
\input{ch06_atmosphere_and_instruments.tex}

%%%%%%%%%%%%%%%%%%%%%%
\input{ch07_airspace_and_flarm.tex}

%%%%%%%%%%%%%%%%%%%%%%
\input{ch08_avionics_and_airframe.tex}

%%%%%%%%%%%%%%%%%%%%%%
\input{ch09_quickstart.tex}

%%%%%%%%%%%%%%%%%%%%%%
\input{ch10_infobox_reference.tex}

%%%%%%%%%%%%%%%%%%%%%%
\input{ch11_configuration.tex}

%%%%%%%%%%%%%%%%%%%%%%
\input{ch12_data_files.tex}

%%%%%%%%%%%%%%%%%%%%%%
\input{ch13_history.tex}

%%%%%%%%%%%%%%%%%%%%%%
\appendix

\chapter{GNU General Public License}\label{cha:gnu-general-public}
\input{gpl.tex}

\end{document}
